### 2021-01-28 appendix.tex
Added scheduler example in the appendix.

>>>>

\section{Milner's Cyclic Scheduler}
\label{sec:milner}
As an example of a deadlock-free cyclic process, Dardha and Gay~\cite{dardhagay18} introduce an implementation of Milner's cyclic scheduler~\cite{milner89} in Priority CP. We reproduce that scheduler here, and show its translation to Priority GV.

\begingroup
\usingnamespace{pcp}
\begin{example}[Milner's Cyclic Scheduler, PCP]
  \label{ex:pcp-scheduler}
  A set of processes $\tm{\proc{i}}$, $1\leq{i}\leq{n}$, is scheduled to perform some tasks in cyclic order, starting with $\tm{\proc1}$, ending with $\tm{\proc{n}}$, and notifying $\tm{\proc1}$ when all processes have finished.

  Our scheduler $\tm{\sched}$ consists of set of agents $\tm{\agent{i}}$, $1\leq{i}\leq{n}$, each representing their respective process. Each process $\tm{\proc{i}}$ waits for the signal to start their task on $\tm{a'_i}$, and signals completion on $\tm{b'_i}$. Each agent signals their process to start on $\tm{a_i}$, waits for their process to finish on $\tm{b_i}$, and then signals for the next agent to continue on $\tm{c_i}$. The agent $\tm{\agent1}$ initiates, then waits for every other process to finish, and signals $\tm{\proc1}$ on $\tm{d}$. Every other agent $\tm{\agent{i}}$, $2\leq{i}\leq{n}$ waits on $\tm{c'_{i-1}}$ for the signal to start. Each of the channels in the scheduler is of a terminated type, and is merely used to synchronise.

  Below is a diagram of our scheduler instantiated with three processes:
  \begin{center}
    \begin{tikzpicture}
      \node[shape=circle,draw=black]                       (agent1) {\tm{\agent1}};
      \node[shape=circle,draw=black,below right=of agent1] (agent2) {\tm{\agent2}};
      \node[shape=circle,draw=black,below left =of agent1] (agent3) {\tm{\agent3}};

      \node[shape=circle,draw=black,above      =of agent1] (proc1) {\tm{\proc1}};
      \node[shape=circle,draw=black,below right=of agent2] (proc2) {\tm{\proc2}};
      \node[shape=circle,draw=black,below left =of agent3] (proc3) {\tm{\proc3}};

      \draw[->]
      (agent1) -- node[pos=0.35,above] {$c_1$}
      node[pos=1.0,above] {$c'_1$} ++ (agent2);
      \draw[->]
      (agent2) -- node[pos=0.1,above] {$c_2$}
      node[pos=0.9,above] {$c'_2$} ++ (agent3);
      \draw[->]
      (agent3) -- node[pos=0.0,above] {$c_3$}
      node[pos=0.65,above] {$c'_3$} ++ (agent1);

      \draw[implies-implies,double]
      (agent1) -- node[pos=0.15,left] {$a_1$} node[pos=0.85,left] {$a'_1$}
      node[pos=0.15,right] {$b_1$} node[pos=0.85,right] {$b'_1$} ++ (proc1);
      \draw[implies-implies,double]
      (agent2) -- node[pos=0.15,above right] {$a_2$} node[pos=0.85,above right] {$a'_2$}
      node[pos=0.15,below left] {$b_2$} node[pos=0.85,below left] {$b'_2$} ++ (proc2);
      \draw[implies-implies,double]
      (agent3) -- node[pos=0.15,above left] {$a_3$} node[pos=0.85,above left] {$a'_3$}
      node[pos=0.15,below right] {$b_3$} node[pos=0.85,below right] {$b'_3$} ++ (proc3);

      \draw[->]
      (agent1) to[bend left=45] node[pos=0.15,left] {$d$} node[pos=0.85,left] {$d'$} (proc1);

      \draw[->,dashed]
      (proc1.east) to[bend left] node[pos=0.15,align=center,above right] {optional\\data transfer} (proc2.north);
      \draw[->,dashed]
      (proc2.west) to[bend left] (proc3.east);
      \draw[->,dashed]
      (proc3.north) to[bend left] (proc1.west);
    \end{tikzpicture}
  \end{center}

    We implement the scheduler as follows, using $\tm{\prod_{I}P_i}$ to denote the parallel composition of the processes $\tm{P_i}$, $\tm{i}\in\tm{I}$, and $\tm{\plug{P}{Q}}$ to denote the plugging of $\tm{Q}$ in the one-hole process-context $\tm{P}$. The process-contexts $\tm{P_i}$ represent the computations performed by each process $\tm{\proc{i}}$. The process-contexts $\tm{Q_i}$ represent any post-processing, and any possible data transfer from $\tm{\proc{i}}$ to $\tm{\proc{i+1}}$. Finally, $\tm{Q_1}$ should contain $\tm{\labwait{d'}}$.
\[
  \begin{array}{lrlr}
    \tm{\sched}
    & \defeq & \tm{\res{a_1}{a'_1}{\dots{\res{a_n}{a'_n}{}}}}
               \tm{\res{b_1}{b'_1}{\dots{\res{b_n}{b'_n}{}}}}
               \tm{\res{c_1}{c'_1}{\dots{\res{c_n}{c'_n}{}}}}
               \tm{\res{d}{d'}{}}
    \\ &     & \tm{(
               \ppar{\proc1}{\agent1}
               \parallel
               \prod_{2\leq{i}\leq{n}}(\ppar{\proc{i}}{\wait{c'_{i-1}}{\agent{i}}})
               )}
    \\
    \tm{\agent1}
    & \defeq & \tm{\close{a_i}{\wait{b_i}{\close{c_i}{\wait{c'_n}{\close{d}{\halt}}}}}}
    \\
    \tm{\agent{i}}
    & \defeq & \tm{\close{a_i}{\wait{b_i}{\close{c_i}{\halt}}}}
    \\
    \tm{\proc{i}}
    & \defeq & \tm{\wait{a'_i}{\plug{P_i}{\close{b'_i}{Q_i}}}}
  \end{array}
\]
\end{example}
\endgroup

\begingroup
\usingnamespace{pgv}
\begin{example}[Milner's Cyclic Scheduler, PGV]
  \label{ex:pgv-scheduler}
  The PGV scheduler has exactly the same behaviour as the PCP version in~\cref{ex:pcp-scheduler}. It is implemented as follows, using $\tm{\prod_{I}\conf{C}_i}$ to denote the parallel composition of the processes $\tm{\conf{C}_i}$, $\tm{i}\in\tm{I}$, and $\tm{\plug{M}{N}}$ to denote the plugging of $\tm{N}$ in the one-hole term-context $\tm{M}$. For simplicity, we let $\tm{\sched}$ be a configuration. The terms $\tm{M_i}$ represent the computations performed by each process $\tm{\proc{i}}$. The terms $\tm{N_i}$ represent any post-processing, and any possible data transfer from $\tm{\proc{i}}$ to $\tm{\proc{i+1}}$. Finally, $\tm{N_1}$ should contain $\tm{\wait\;{d'}}$.
  \[
    \begin{array}{lrlr}
      \tm{\sched}
      & \defeq & \tm{\res{a_1}{a'_1}{\dots{\res{a_n}{a'_n}{}}}}
                 \tm{\res{b_1}{b'_1}{\dots{\res{b_n}{b'_n}{}}}}
                 \tm{\res{c_1}{c'_1}{\dots{\res{c_n}{c'_n}{}}}}
                 \tm{\res{d}{d'}{}}
      \\ &     & \tm{\begin{array}{lll}
                       (
                       & \phi\;\proc1
                         \parallel
                         \child\;\andthen{\agent1}{\andthen{\wait\;{c'_n}}{\close\;{d}}}
                       \\
                       \parallel
                       &
                         \prod_{2\leq{i}\leq{n}}
                         (\child\;\proc{i} \parallel \child\;\andthen{\wait\;{c'_{i-1}}}{\agent{i}})
                       & )
                     \end{array}}
      \\
      \tm{\agent{i}}
      & \defeq & \tm{\andthen{\close\;{a_i}}{\andthen{\wait\;{b_i}}{\close\;{c_i}}}}
      \\
      \tm{\proc{i}}
      & \defeq & \tm{\andthen{\wait\;{a'_i}}{\plug{M_i}{\andthen{\close\;{b'_i}}{N_i}}}}
    \end{array}
  \]
\end{example}
\endgroup

If $\pcp{\tm{\cpgvM{P_i}}}=\pgv{\tm{M_i}}$ and $\pcp{\tm{\cpgvM{Q_i}}}=\pgv{\tm{N_i}}$, then the translation of $\pcp{\tm{\sched}}$ (\cref{ex:pcp-scheduler}), $\pcp{\tm{\cpgvC{\sched\,}}}$, is exactly $\pgv{\tm{\sched}}$ (\cref{ex:pgv-scheduler}).
